### 2023-12-24 sections/4-results.tex
justification of random lfd

In contrast with Figure~\ref{fig:results:sift-lfd}, Figure~\ref{fig:results:random-lfd} shows the LFD by depth for a random dataset with the same cardinality and dimensionality as sift. 
We observe that the LFD starts as high as 20 (for all data) at depth 0. As shown by the needlepoint shape of the plot, the LFD for all clusters 
is about the same from depth 0 through depth 5. 
After depth 5, we begin to see a wider spread between the maximum and minimum LFD at each depth. 
The median decreases seemingly linearly from about 20 at depth 0 to about 2 at depth 20, remaining less than 2 for all depths greater than 20.

>>>>

In contrast with Figure~\ref{fig:results:sift-lfd}, Figure~\ref{fig:results:random-lfd} shows the LFD by depth for a random dataset with the same cardinality and dimensionality as sift. 
We observe that the LFD starts as high as 20 (for all data) at depth 0. As shown by the needlepoint shape of the plot, the LFD for all clusters 
is about the same from depth 0 through depth 5. 
After depth 5, we begin to see a wider spread between the maximum and minimum LFD at each depth. 
The median decreases seemingly linearly from about 20 at depth 0 to about 2 at depth 20, remaining less than 2 for all depths greater than 20.
The LFD of approximately 20 at depth 0 is what we expect for a random dataset.
To elaborate, the distribution of points in such a dataset should reflect the curse of dimensionality, or the fact that in high dimensional spaces, the minimum pairwise distance between any two points is approximately equal to maximum pairwise distance. 
As a result, the root cluster's radius $r$, which reflects the maximum distance between the center $c$ and any other point, should not differ significantly from the distance between the center and its closest point. 
A consequence of this is that with high probability, a point in the root cluster is a distance greater than half the root radius from the center; $B_X(c, \tfrac{r}{2})$ contains only $c$ while $B_X(c, r) = |X|$, where $X$ is the dataset. 
Given our definition of LFD in Equation~\ref{eq:methods:lfd-simplified}, this means that the LFD of the root cluster is approximately $\log_2(\frac{|X|}{1}) = \log_2(1,000,000) \approx 20$, which is what we observe in Figure~\ref{fig:results:random-lfd}.
